%!TEX root = thesis.tex
%%%%%%%%%%%%%%%%%%%%%
\chapter*{Acknowledgments}
\addcontentsline{toc}{chapter}{Acknowledgments}
Now that I am right at the finish line, besides being happy that the exhaustive part is finally over, I feel a great deal of gratitude.
I would simply not have made it through, were it not for the wonderful people that take part in my life.

First of all I want to thank the members of the examination board for taking the time to evaluate the dissertation.
I know that you are all incredibly busy, and truly appreciate your efforts in training the next generation of scientists.

I want to express my deep gratitude towards {\it Prof. Michele Burrello}.
You are one of the first persons I approach for professional advice, and I am always astonished by how much you know about physics.
Thank you so much for your mentoring, and for inviting me to the Niels Bohr Institute.
Brainstorming with smart and inspiring people is one of the best experiences during my PhD.
One of the best example out of of this ``equivalence class'' of individuals is {\it Prof. Pietro Massignan}.
I am looking forward to visiting the post-pandemic Primavera Sound Festival with you and Hélia.
Most of my academic collaborators are Italian, and although I am not a fan of cliches, I must admit that they are very passionate in general:
Thank you, {\it Michele Filippone}, for your regular motivation speeches during the time we worked together, for having me around in Geneva, for showing me the Bains des Pâquis and warning me about drinking cold water after a Swiss Fondue.
Speaking of passionate Italians, I want to express my gratitude towards {\it Jun.-Prof. Jamir Marino}.
Thank you ``für den frischen Wind'' as we Germans like to say, for your feedback regarding my thesis and for putting so much effort in a lively scientific exchange, which is incredibly hard during this pandemic.
By now it is clear to me that an academic career is unavoidably linked to a level of uncertainty: be it compensation, doubting yourself or feeling unrecognized and irrelevant, it results in a lot of unnecessary distractions from the actual work.
I am grateful to the head of our group, {\it Prof. Peter van Dongen}, who managed to eliminate most of these uncertainties during the period of my PhD.
Thank you for your scientific advice, for being part of the examination board, and for carefully reading this thesis.
I am glad that you consider me part of your little ``bubble'', and I wish you the best for the future.
I want to thank {\it Prof. Patrick Windpassinger} for being part of the examination board and for writing a letter of recommendation to the graduate school.
It has always been a pleasure to chat with you when we met at the institute's staircase.
My PhD would have been impossible without the scholarship of the Graduate School Materials Science in Mainz (MAINZ) and the Max Planck Graduate Center (MPGC).
Special thanks to {\it Michael Fuchs}, {\it Katrin Klauer}, and to {\it Prof. Thomas Speck}.

A student's life would be pretty sad and lonely without a group of fellows:
Special thanks to my dear friends and colleagues, {\it Matteo Campo}, {\it Álvaro Díaz Fernández}, {\it Achim Harzheim}, {\it William Janke}, {\it Johannes Jünemann}, {\it Philipp Schmoll} and {\it Niklas Tausendpfund}, for taking the journey with me, and for reading part of this thesis.
I will miss our joint conferences, lunches and coffee breaks.

Music is a big part of my life, and it hurts me very much that these past few months were exceptionally quiet.
To all my musician friends, in particular to {\it Simon Engelhardt}, {\it Carlos Wagner} and the Grundfunk family:
There will be a time when we enjoy once more the incredible joy of playing with real people, at a real stage, and in front of an audience.
Until then, hold on tight and keep on practicing.

Sometimes (mostly Tuesdays) I came in late due to the merciless morning squash matches between me and {\it Andreas Dittinger}, {\it Fabian Kolf} and {\it Maximilian Shaikh-Yousef}.
Before my passion for squash, I was happily climbing like an ape during my free evenings in the Blockwerk Mainz with {\it Johannes} and {\it Julia Hofmann}, {\it Ruben Hoffmann} and the BFE crew, which again resulted in me coming in late to university.
At other times, I overslept because of a merry wine-drinking session with our favorite upstairs neighbors {\it Laura Hollegger} and {\it Max Henderson}.
To all of my dear friends: Thanks for the joy and happiness you bring into my life.

Family always comes first, although by now we live a bit detached from each other.
I want to thank my brother-in-law, {\it Stefan Stamer}, and my parents-in-law, {\it Helma Heinz-Stamer} and {\it Thomas Stamer}.
You know me for half of my life now, and I can only imagine how annoying it must have been for you to meet me during my teens, hitting on your sister/daughter.
Joking aside, I always feel warm and welcome when I am around you, and I thank you deeply for trusting me with the biggest treasure you have:
{\it Susanne}, you are the love of my life.
I don't know if you remember our chat nine years ago, but you are the reason why I pursued a career as a Physicist, and I will never forget that.
I want to thank my grandma, {\it Irmgard Heinz}, and I hope this message reaches you somehow.
You were the sweetest person of age that I know, always happy when we passed by, and never showed your disappointed if we were busy otherwise.
I remember our joint trip to the canary islands with Susanne like it was yesterday, because it woke my curiosity to travel the world.

To my siblings, {\it Michael Haller} and {\it Astrid Fengler}.
Thank you for your patience with me.
% I've not always been the best little brother, and certainly was a rascal when I was younger.
Despite all our differences, I know that I can always count on you, and I am deeply grateful for having you in my life.
% Since my parents do not understand english, but certainly deserve one of the biggest thanks, allow me to switch languages for a while.
{\it Liebe Mama, lieber Papa}, ihr habt meine Entscheidung Physik zu studieren stets unterstützt, auch wenn ihr keinen direkten Bezug zu dem Thema habt.
Euch ist es stets das Wichtigste, dass eure Kinder eigene Entscheidungen treffen und diese dann auch realisieren.
Nur durch dieses große Vertrauen in uns sind wir zu den Personen geworden, die wir heute sind.
Dafür liebe ich euch und bin euch unendlich dankbar.

Al mio capo {\it Prof. Matteo Rizzi}.
You made all of this possible by shaping me into the scientist I am today.
I remember very well when I entered your office for the first time, not knowing what to expect, a little shy and intimidated by your wisdom.
Although I knew so little at the time we started working together, you never lost interest and aided me to the best of your abilities.
I could not have been guided better and I am proud to call myself your student.
With all of my heart, I deeply thank you for our joint time and wish the best to you, Michela, and Giovanni.

I am grateful to everyone that shaped and decorated the path of my PhD.
It has been a wonderful experience, and I would do it all over again.
